### 2022-08-08 report/rl_qf_report.tex
make comparison for three algorithms

\documentclass[11pt]{article}
\usepackage{amsmath, amsfonts, bm, listings, enumitem, geometry, titlesec,
array, makecell, caption, graphicx, float, algorithm, algpseudocode, siunitx}
\usepackage[table]{xcolor}
\usepackage{tikz}
\usetikzlibrary{shapes.geometric, arrows.meta}
\usepackage{hyperref}
\geometry{margin=0.8in}
\graphicspath{ {../data/} }

%\renewcommand{\familydefault}{\sfdefault}
%\sisetup{output-exponent-marker=\ensuremath{\mathrm{e}}}
\tikzstyle{process} = [rectangle, minimum width=2cm, minimum height=1.3cm, 
text centered, draw=black]
\tikzstyle{arrow} = [thick,->,>={Stealth[scale=1.1]}]

\title{Reinforcement Learning Report}
\author{Yuanhao JIANG}
\date{}

\begin{document}
\maketitle
\tableofcontents

\section{Introduction}
The object is to use reinforcement learning to solve the following scenario:
a customer arrives with a number of features, we want to promote the customer
with a proper price, such that he is most likely to purchase our product, and 
we also have the highest profit and a better portfolio. 
\vspace{3mm}\newline We can consider this scenario as markov decision process,
with the current state being the current customer features and current 
portfolio. At each time step, the agent promote a price according to the 
current customer features and the current portfolio, after that the customer 
resopnse to the action. The reward is then given besed on the customer 
response and the portfolio. Then the portfolio updates and a new customer 
arrives.

\section{Reinforcement Learning}
The reinforcement learning method we want to use is based on the Markov 
decision process (MDP), as illustrated below:
\begin{center}
    \begin{tikzpicture}
        \node (agent) [process, minimum width=2cm, minimum height=2cm,
            label={[anchor=north,inner sep=5pt]north:Agent}]{};
        \node (pi) [process, left of=agent, minimum width=1cm, 
            minimum height=1cm,xshift=1cm, yshift=-0.3cm] {\(\pi\)};
        \node (env) [process, minimum height=1.5cm, below of=pi, yshift=-1cm] 
            {Environment};

        \draw [arrow] (env.163) -- ++(-1.5,0) node[anchor=south, xshift=0.8cm]
            {\(S_{t+1}\)} coordinate(a1);
        \draw [arrow] (a1) -- ++(-0.7,0) |- node[anchor=south, xshift=1.3cm] 
            {\(S_t\)} (pi);
        \draw [arrow] (env.197) -- ++(-1.5,0) node[anchor=south, xshift=0.8cm]
            {\(R_{t+1}\)} coordinate(a2);
        \draw [arrow] (a2) -- ++(-1.3,0) -- ++(0,3.2) 
            -- node[anchor=south, xshift=0.5cm] {\(R_t\)} (agent.154);
        \draw [arrow] (pi) -- ++(2.7,0) node[anchor=south, xshift=-0.8cm] 
            {\(A_t\)} |- (env);
        \draw [dashed,thick] (-2.7,-1.5) -- (-2.7, -3.2);
    \end{tikzpicture}

\end{center}
\noindent However, in RL, the agent is not told which actions to take, but 
instead must discover which actions yield the most reward by trying them. Our 
goal at each time step is to find a policy \(\pi\) to maximize the expected 
reward alonge the process afterwards, which is given by the value function
\[V(s)=\mathbb{E}\left[\sum_{i=0}^{\infty} \gamma^i R_{t+i+1}\;\middle|\; 
S_t=s\right]\]
To achieve this, we need to evaluate the current policy, and improve it 
accordingly.

\subsection{Policy Evaluation}
We want to evaluate the value function under a given policy \(\pi\). According 
to Bellman equation,
\[v_\pi(s)=\mathbb{E}_\pi\left[R_{t+1}+\gamma v_\pi(S_{t+1})\;\middle|\;
S_t=s\right]\]
We can use it to iteratively compute \(v_\pi\). The iteratively update is
\[v_{k+1}(s)=\mathbb{E}_\pi\left[R_{t+1}+\gamma v_k(S_{t+1})\;\middle|\;
S_t=s\right]\]
which is proven that it will converge to the true value given any initial value 
function.

\subsection{Policy Improvement}
Now we have the evaluated value function for the policy, we want to improve it
accordingly. To do this, we introduce the q-function under policy \(\pi\)
\[q_\pi(s, a) = \mathbb{E}\left[R_{t+1} + \gamma v_\pi(S_{t+1})\;\middle|\;
S_t=s, A_t=a\right]\]
The policy improvement theorem says that, let \(\pi\) and \(\pi'\) be any pair
of policy, for all \(s\) we have
\[q_\pi(s, \pi'(s))\geq v_\pi(s) \:\Rightarrow\: v_{\pi'}(s) \geq v_\pi(s)\]
Now we can improve the policy by
\[\pi'(s)=\underset{a}{\text{argmax}}\:q_\pi(s,a)\]
for all \(s\).

\subsection{Policy Iteration}
By iteratively applying policy evaluation and policy improvement, we can obtain
a sequence of monotonically improving policies and value functions:
\[\pi_0 \overset{E}{\longrightarrow} v_{\pi_0} \overset{I}{\longrightarrow}
\pi_1 \overset{E}{\longrightarrow} v_{\pi_1} \overset{I}{\longrightarrow}
\dots \overset{I}{\longrightarrow} \pi_* \overset{E}{\longrightarrow} v_*\]
where \(\overset{E}{\longrightarrow}\) denotes a policy evaluation and
\(\overset{I}{\longrightarrow}\) denotes a policy improvement. This way of 
finding an optimal policy is called policy iteration.

\subsection{Value Iteration}
One drawback to policy iteration is that each of its iterations involves policy
evaluation, which itself is a protracted iterative computation requiring 
multiple sweeps through the state set.
\vspace{3mm}\newline We now truncate the policy evaluation such that it is
stopped after just one sweep (one update of each state):
\begin{align*}
    v_{k+1}(s)&=\underset{a}{\text{max}}\:q(s, a)\\
              &=\underset{a}{\text{max}}\:\mathbb{E}\left[
          R_{t+1}+\gamma v_k(S_{t+1})\;\middle|\;S_t=s, A_t=a\right]
\end{align*}
We keep iterate until \(v\) converges (to \(v_*\)), then we record the last
policy \(\pi(s) = \underset{a}{\text{argmax}}\:q(s, a)\).

\subsection{Generalized Policy Iteration}
As long as both processes, policy evaluation and policy improvement, continue
to update all states, the ultimate result is typically the same --- converge to
the optimal value function and an optimal policy.
\vspace{3mm}\newline Generalized policy iteration (GPI) is used to refer to the
general idea of letting policy evaluation and policy improvement interact, 
independent of the granularity of the two processes (instead of letting each
complete before the other begins).

\subsection{Temporal-Difference Learning}
When we don't have the complete model of the environment, we need to estimate 
the value function from only experience. There are two methods to do so: 
Monte Carlo Methods and Temporal-Difference. We will use Temporal-Difference
method in this project.
\vspace{3mm}\newline To estimate \(v_\pi(S_t)\) using only the experience, we 
take the average on all experienced trajectory' values. Temporal-Difference
method uses \(G_i = [R_{t+1}]_i+\gamma [v_\pi(S_{t+1})]_i\) to estimate the 
value of the \(i\)th trajectory starting at \(S_t\), with \(v_\pi\) being the 
estimated value function at the \(i\)th time \(S_t\) is visitied, which can be 
initialized to anything at start. Now we have the incremental formula to update
our value function:
\begin{align*}
    [v_\pi(S_t)]_{k+1} &= \frac{1}{k}\sum_{i=1}^{k} G_i\\
               &= \frac{1}{k}\left(G_k + \sum_{i=1}^{k-1} G_i\right)\\
               &= \frac{1}{k}\left(G_k + (k-1)[v_\pi(S_t)]_{k}\right)\\
               &= [v_\pi(S_t)]_k + \frac{1}{k}(G_k-[v_\pi(S_t)]_k)\\
               &= [v_\pi(S_t)]_k + \frac{1}{k}
               ([R_{t+1}]_k+\gamma [v_\pi(S_{t+1})]_k-[v_\pi(S_t)]_k)
\end{align*}
To simplify:
\[v_\pi(S_t)\longleftarrow v_\pi(S_t)+\alpha
[R_{t+1}+\gamma v_\pi(S_{t+1})-v_\pi(S_t)]\]
with \(\alpha\) being the stepsize. And this also applies to the q function:
\[q_\pi(S_t, A_t)\longleftarrow q_\pi(S_t, A_t)+\alpha
[R_{t+1}+\gamma q_\pi(S_{t+1}, A_{t+1})-q_\pi(S_t, A_t)]\]

\subsection{Policy Gradient Methods}
When the state space is arbitrarily large, we cannot find an optimal policy or
the optimal value function given the limited resources and time. Instead of 
keeping track of the value and the action(s) to select for each state (tabular 
methods), we can use parameterized functions, for example, artificial neural 
networks (ANNs), to approximate the value function and the policy function. 
\vspace{3mm}\newline Previously methods learned the values of actions and then 
select actions based on their estimated action values (q function). Now we 
consider methods that instead learn a parameterized policy that select actions 
without consulting a value function. A value function may still be used to 
learn the policy parameter, but is not required for action selection.
\vspace{3mm}\newline We use \(\bm{\theta}\) and \(\bm{w}\) to denotes the 
policy's parameter vector and value function's weight vector, respectively.
Then we have our policy \(\pi\left(a\;\middle|\;s, \bm{\theta}\right)=
\text{Pr}\left\{A_t=a\;\middle|\;S_t=s, \bm{\theta_t}=\bm{\theta}\right\}\) and
value function \(v_\pi(s, \bm{w})\). We also define the scalar performance 
measure (objective) to be 
\[J(\bm{\theta})=\begin{cases}
    \ v_{\pi_{\bm{\theta}}}(s_0); &\text{episodic case}\\
    \ \lim_{t \to \infty} \mathbb{E}
    \left[R_t\;\middle|\;S_0,A_{0:t-1}\sim\pi\right];      
                                  & \text{continuing case}
\end{cases}\]
Policy gradient methods seeks to maximize this performance measure, so their 
updates approximate gradient ascent in \(J\):
\[\bm{\theta}_{t+1}=\bm{\theta}_t+\alpha\widehat{\nabla J(\bm{\theta}_t)}\]
where \(\widehat{\nabla J(\bm{\theta}_t)}\) is a stochastic estimate whose
expection approximates the gradient of the performance measure w.r.t 
\(\bm{\theta}_t\).

\subsubsection{The Policy Gradient Theorem}
The policy gradient theorem says
\[\nabla J(\bm{\theta})\propto \sum_{s} \mu(s) 
\sum_{a} q_\pi(s,a)\nabla\pi\left(a\middle|s, \bm{\theta}\right)\]
where \(\mu\) is the stationary distribution of the state. Thus we can derive
\begin{align*}
    \nabla J(\bm{\theta}) 
    &\propto \sum_{s} \mu(s) 
    \sum_{a} q_\pi(s,a)\nabla\pi\left(a\middle|s, \bm{\theta}\right)\\
    &= \mathbb{E}_\pi\left[\sum_{a} q_\pi(S_t,a)
    \nabla\pi\left(a\middle|S_t, \bm{\theta}\right)\right]\\
    &= \mathbb{E}_\pi\left[\sum_{a} \pi\left(a\middle|S_t, \bm{\theta}\right)
        q_\pi(S_t,a)\frac{\nabla\pi\left(a\middle|S_t, \bm{\theta}\right)}
        {\pi\left(a\middle|S_t, \bm{\theta}\right)}\right]\\
    &=\mathbb{E}_\pi\left[q_\pi(S_t,A_t)
        \frac{\nabla\pi\left(A_t\middle|S_t, \bm{\theta}\right)}
        {\pi\left(A_t\middle|S_t, \bm{\theta}\right)}\right]
\end{align*}
The policy gradient theorem can be generalized to include a comparison of the 
action value to an arbitrary baseline \(b(s)\):
\[\nabla J(\bm{\theta})\propto \sum_{s} \mu(s) \sum_{a} 
    \Bigl(q_\pi(s,a)-b(s)\Bigr)\nabla\pi\left(a\middle|s, \bm{\theta}\right)\]
with the same derivation we have
\[\nabla J(\bm{\theta})\propto\mathbb{E}_\pi\left[
        \Bigl(q_\pi(S_t,A_t)-b(S_t)\Big)
        \frac{\nabla\pi\left(A_t\middle|S_t, \bm{\theta}\right)}
{\pi\left(A_t\middle|S_t, \bm{\theta}\right)}\right]\]
And the discounted version of above is:
\[\nabla J(\bm{\theta})\propto\mathbb{E}_\pi\left[
        \gamma^{t}\Bigl(q_\pi(S_t,A_t)-b(S_t)\Big)
        \frac{\nabla\pi\left(A_t\middle|S_t, \bm{\theta}\right)}
{\pi\left(A_t\middle|S_t, \bm{\theta}\right)}\right]\]
One natural choice for baseline is an estimate of the state value, 
\(\hat{v}(S_t, \bm{w})\), where \(\bm{w}\) is the weight vector to be learned.

\subsubsection{Actor-Critic Methods}
Recall the TD method, we make use of 
\(q_\pi(s,a)=\mathbb{E}\left[R_{t+1}+\gamma v_\pi(S_{t+1})\;\middle|\;
S_t=s, A_t=a\right]\), by taking one sample estimate, and choosing
\(\hat{v}(S_t, \bm{w})\) as baseline, we have
\[\nabla J(\bm{\theta})\propto\mathbb{E}_\pi\left[
        \Bigl(R_{t+1}+\gamma \hat{v}(S_{t+1}, \bm{w})-\hat{v}(S_t, \bm{w})\Big)
        \frac{\nabla\pi\left(A_t\middle|S_t, \bm{\theta}\right)}
{\pi\left(A_t\middle|S_t, \bm{\theta}\right)}\right]\]
Again, by taking one sample estimate we have the update formula for 
\(\bm{\theta}\):
\begin{align*}
    \bm{\theta}_{t+1}
    &=\bm{\theta}_t + \alpha_{\bm{\theta}}
    \Bigl(R_{t+1}+\gamma \hat{v}(S_{t+1}, \bm{w})-\hat{v}(S_t, \bm{w})\Big)
    \frac{\nabla\pi\left(A_t\middle|S_t, \bm{\theta}_t\right)}
    {\pi\left(A_t\middle|S_t, \bm{\theta}_t\right)}\\
    &=\bm{\theta}_t+\alpha_{\bm{\theta}}\delta_t
    \nabla\ln{\pi\left(A_t\middle|S_t, \bm{\theta}_t\right)}
\end{align*}
As for the weight vector \(\bm{w}\) for the estimated value function, at each 
time step \(t\), we want to update it by minimizing the squared error,
\(\bigl[v_\pi(S_t)-\hat{v}(S_t, \bm{w}_t)\bigr]^2\):
\begin{align*}
    \bm{w}_{t+1} 
    &= \bm{w}_t - 
    \alpha_{\bm{w}}'\nabla \Bigl[v_\pi(S_t)-\hat{v}(S_t, \bm{w}_t)\Bigr]^2\\
    &= \bm{w}_t - \alpha_{\bm{w}}\Bigl(v_\pi(S_t)-\hat{v}(S_t, \bm{w}_t)
    \Bigr)\nabla \hat{v}(S_t, \bm{w}_t)
\end{align*}
where \(v_\pi\) is the true value function. Again, since \(v_\pi(s)=
\mathbb{E}\left[R_{t+1}+\gamma v_\pi(S_{t+1})\;\middle|\;S_t=s\right]\), we
take one sample estimate, which gives us the update formula for \(\bm{w}\):
\begin{align*}
    \bm{w}_{t+1} 
    &= \bm{w}_t - \alpha_{\bm{w}}
    \Bigl(R_{t+1}+\gamma\hat{v}(S_{t+1},\bm{w})-\hat{v}(S_t, \bm{w}_t)
    \Bigr)\nabla \hat{v}(S_t, \bm{w}_t)\\
    &=\bm{w}_t - \alpha_{\bm{w}}\delta_t\nabla \hat{v}(S_t, \bm{w}_t)
\end{align*}
The actor-critic method is based on the above updates formulas:
\begin{algorithm}[H]
\caption{Actor-Critic}
\begin{algorithmic}[1]
\State Input: a differentiable policy parameterization 
\(\pi\left(a\middle|s, \bm{\theta}\right)\)
\State Input: a differentiable state-value function parameterization
\(\hat{v}(s,\bm{w})\)
\State Step sizes \(\alpha_{\bm{\theta}}>0\), \(\alpha_{\bm{w}}>0\)
\State \(\bm{\theta} \gets \bm{0}\), \(\bm{w} \gets \bm{0}\)
\Loop \hspace{0.3mm} (for each episode)
    \State Initialize \(S\) (first state of the episode)
    \Loop \hspace{0.3mm} (for each time step \(i\))
        \State \(A\sim\pi\left(\cdot\middle|S, \bm{\theta}\right)\)
        \State Take action \(A\), observe \(S'\), \(R\)
        \State \(\delta \gets R + \gamma\hat{v}(S',\bm{w})-\hat{v}(S,\bm{w})\)
        \State \(\bm{w} \gets \bm{w} + 
        \alpha_{\bm{w}} \delta\nabla\hat{v}(S,\bm{w})\)
        \State \(\bm{\theta} \gets \bm{\theta} 
        + \alpha_{\bm{\theta}}\gamma^{i}\delta
        \nabla\ln{\pi\left(A\middle|S, \bm{\theta}\right)}\)
        \State \(S \gets S'\)
    \EndLoop
\EndLoop
\end{algorithmic}
\end{algorithm}

\section{Build the Environment}
The environment does the following things:
\begin{enumerate}
    \item record the current state that can be queried anytime
    \item when an action, i.e. a price, is given according to the current 
        state, it outputs the reward, and goes to next state
\end{enumerate}
At each time step, the state consists of a randomly generated customer's 
features and the current portfolio. After the agent performs an action, i.e.
promotes a price to the customer, the reward is then computed based on the 
profit on this customer and the portfolio up to now.

\subsection{Customer Features}
A customer' features is represented by an array (or vector), denoted by \(x\) 
with 16 entries with each generated randomly from a distribution, 
including normal distribution, binomial distribution, and so on. For example, 
a feature vector \(x\) is described as below:
\begin{table}[htpb]
    \centering
    \begin{tabular}{c|c|c|c|c|c|c|c}
        gender & age & car cost & miles & brand & random feature 0 & random 
        feature 1 & \dots\\
        \hline
        1.0 & 42.723 & 94032.096 & 30086.311 & 29.264 & 8.0 & 68.740 & \dots
    \end{tabular}
\end{table}

\noindent After a price, denoted by \(c\), is promoted to a customer, the 
customer will give a resopnse, denoted by \(r\), which is 0 if he does not buy
the product, and 1 otherwise. And the profit we have for that customer, 
denoted by pf, can be computed by
\[\text{pf} = r*\left(c - g(x)\right)\]
where \(g(x)\) is the cost on that customer.

\subsection{Resopnse}
The resopnse is generated by a generalized linear model, whose input is a 
vector of a customer features vector \(x\) appended with a price \(c\), i.e.
\(r=glm(x[0], \dots, x[14], c)\).
\vspace{3mm}\newline The GLM used in the project is fitted using a randomly
generated data set with each observation is a customer's features followed by 
a price, then followed by a resopnse e.g.
\begin{table}[htpb]
    \centering
    \begin{tabular}{c|c|c|c|c|c}
        & customer feature 0 & \dots & customer feature 15 & price & resopnse\\
        \hline
        1 & 1.0 & \dots & 2.0 & 1109.504 & 1.0\\
        \hline
        2 & 0.0 & \dots & 1.0 & 665.8 & 0.0\\
        \hline
        \dots & \dots & \dots & \dots & \dots & \dots
    \end{tabular}
\end{table}

\noindent And its link function is chose to be the logistic function.

\subsection{State}
The state at start of each time step is the feature vector of the new customer 
to promote (might be modified a bit), appended by the portfolio, which is also 
a data vector consists of the following entries:
\begin{itemize}
    \item average profit over all customers so far, expect for the new customer
        at current time step, denoted by avg\_pf (initialized to \(0\)), 
        which is computed by
        \[\text{avg\_pf}_t = \frac{1}{t}\sum_{i=0}^{t-1} \text{pf}_i, \:t>0\]
    \item the portion of the buyers over all customers, expect for the new one,
        denoted by p (initialized to 0), which is computed by
        \[\text{p}_t=\frac{1}{t}\sum_{i=0}^{t-1} r_i, \:t>0\]
    \item the variance of the portion of buyers in their categories, denoted 
        by var: we divide customers (expect for the new one at current time 
        step) into 4 categories, according to their features, and for each 
        category, we compute the portion of the customers who buy our product,
        denoted by \(\text{pt}_i\) with \(i=1, 2, 3, 4\), then we compute the 
        variance of the 4 portions:
        \[\text{var}=\text{Var}(\text{pt})=\frac{\sum_{i=1}^{4} 
        \left(\text{pt}_i-\bar{\text{pt}}\right)^2}{3}\]
        Later we use this to compute reward to 
        make sure the agent will promote to all types of customers instead of 
        only focusing on one type.
        \vspace{2mm}\newline Note that \(\text{var}\leq1\).
\end{itemize}
In the end, the state \(s\), or \(s_t\) is of the form 
\(s_t=(x_t[0],\:\dots,\:x_t[14],\:\text{avg\_pf}_t,\:\text{p}_t,
\:\text{var}_t)\)

\subsection{Reward}
The reward at each time step \(t\), after the price \(c_t\) is promoted to the 
customer \(x_t\), denoted by \(R_t\), is calculated by
\[R_t = \text{pf}_t\left(1-h(\text{var}_t)\right)\]
where \(h\) is a function that indicates how much we care about the variance,
i.e. how important it is to try to promote to customers of all categories. In
this project I choose \(h\) to be such that 
\(h(\text{var}) = \sqrt{\text{var}}/2\).

\section{Build the model}
With the environment well built, we now build the model for our policy and the
value function. In this project we use artificial neuro networks (ANNs) as the 
parameterized policy and value function, each with three linear layers, and 
with the ReLU function as the activation function. And the input of both 
function is the state, as illustrated below:
\[S_t\rightarrow\underset{\text{ANN for value function}}{
        \boxed{
            \underset{n(S)\rightarrow 128}{\text{Linear}}
            \rightarrow\text{ReLU}
            \rightarrow\underset{128\rightarrow 128}{\text{Linear}}
            \rightarrow\text{ReLU}
            \rightarrow\underset{128\rightarrow 1}{\text{Linear}}
        }
}\rightarrow v_t\]
\[S_t\rightarrow\underset{\text{ANN for policy function}}{
        \boxed{
            \underset{n(S)\rightarrow 128}{\text{Linear}}
            \rightarrow\text{ReLU}
            \rightarrow\underset{128\rightarrow 128}{\text{Linear}}
            \rightarrow\text{ReLU}
            \rightarrow\underset{128\rightarrow n(A)}{\text{Linear}}
            \rightarrow\underset{(0.1, \: 0.1)}{\text{Threshold}}
        }
}\rightarrow \pi_t\rightarrow c_t\]
where \(n(S)\) denotes the vector length of the state and \(n(A)\) denotes 
the number of actions. 
\vspace{3mm}\newline We generate the policy distribution by scoring each 
action. E.g. denoted the policy function output by \(y\), then the probability 
of choosing action \(i\) (\(i=0, 1, ..., n(A) - 1\)) is computed by 
\[P(i)=\frac{y_i}{\sum_{j=0}^{n(A)-1} y_j}\]
Note that we add a Threshold layer at the end of policy 
function, which is defined by 
\[\text{Threshold}(0.1, 0.1)=\text{max}(0.1, x)\]
which makes sure that 
\begin{enumerate}
    \item the probabilities are valid (no negative values, greater score 
        results in greater probability);
    \item all actions have the probability to be chosen at early stage, which 
        is beneficial for training since it increase the exploration for 
        different actions.
\end{enumerate}

\section{RL with Actor-Critic}
\subsection{The Algorithm}
In this project, we consider the episodic learning procedure, where each 
episode has finite time steps, \(T\), to be specified as a constant, and we 
will also reset the environment at the start of each episode. The algorithm 
below is to use RL, with actor-critic method, and with the help of what we 
have built in previous sections, to solve the problem:
\begin{algorithm}[H]
\caption{Actor-Critic for our financial problem}
\begin{algorithmic}[1]
\State Initialize ANN policy parameterization 
\(\pi\left(c\middle|s, \bm{\theta}\right)\), with any \(\bm{\theta}\)
\State Initialize ANN state-value function parameterization
\(\hat{v}(s,\bm{w})\) with any \(\bm{w}\).
\State Step sizes \(\alpha_{\bm{\theta}}>0\), \(\alpha_{\bm{w}}>0\)
\Loop \hspace{0.3mm} (for each episode)
    \State Reset the environment
    \State Generate an \(x\) (first customer of the episode)
    \State Initialize \(S\) (first state of the episode)
    \For{\(t=0, 1, ..., T\)}
        \State \(c\sim\pi\left(\cdot\middle|S, \bm{\theta}\right)\)
        \State Take action \(c\), observe \(S'\), \(R\) from environment
        \State \(\delta \gets R + \gamma\hat{v}(S',\bm{w})-\hat{v}(S,\bm{w})\)
        \State \(\bm{w} \gets \bm{w} + 
        \alpha_{\bm{w}} \delta\nabla\hat{v}(S,\bm{w})\)
        \State \(\bm{\theta} \gets \bm{\theta} 
        + \alpha_{\bm{\theta}}\gamma^{t}\delta
        \nabla\ln{\pi\left(A\middle|S, \bm{\theta}\right)}\)
        \State \(S \gets S'\)
    \EndFor
\EndLoop
\end{algorithmic}
\end{algorithm}
\noindent The following graph illustrates the algorithm:
\vspace{5mm}
\begin{center}
    \begin{tikzpicture}[node distance=2cm]
        \node (agent) [process, minimum width=6cm, minimum height=3cm,
            label={[anchor=north east,inner sep=7pt]north east:Agent}]{};
        \node (pi) [process, left of=agent, xshift=0.5cm, yshift=-0.3cm] 
            {\(\pi\left(c\middle|s,\bm{\theta}\right)\)};
        \node (v) [process, right of=pi, xshift=1cm] 
            {\(\hat{v}\left(s,\bm{w}\right)\)};
        \node (env) [process, below of=pi, yshift=-2.5cm] {Environment};

        \draw [arrow] (pi) -- node[anchor=east] {c} (env);
        \draw [arrow] (env) 
            -- ++(0,-2.3) node[anchor=east, yshift=0.8cm] {\(S_{t+1}\)}
            -- ++(-1.5,0) coordinate(next_s);
        \draw [arrow] (next_s) -- ++(-1,0) -- ++ (0,10) -- ++(4,0) 
            -- ++(0,-1.7) 
            node[anchor=east, yshift=0.9cm] {\(S_t\)} coordinate(m1)
            -- (pi.93);
        \draw [arrow] (m1) -- (v.93);
        \draw [arrow] (env) 
            -- ++(3,0) coordinate(m2) node[anchor=south, xshift=-1cm] {\(R_t\)} 
            -- +(1.7,0) 
            node[anchor=south, xshift=-0.8cm] {\(\delta\)} coordinate(m3);
        \draw [arrow] (v) -- (m2) node[anchor=west, yshift=1.7cm] {\(v\)};
        \draw [dashed,arrow] (m3) -- ++(0,2.2) coordinate(m4) -- (pi.327);
        \draw [dashed,arrow] (m4) -- (v);
        \draw [dashed, thick] (-3,-6.5) -- (-3,-7.8);
    \end{tikzpicture}
\end{center}

\subsection{Training Result}
The following graph shows the training result of the above algorithm (300
moving average reward vs iteration), with 600 iterations (5 episodes per 
iteration), \(T=300\), \(\gamma=0.99\), 
\(\alpha_\theta=\alpha_{\bm{w}}=\num{3e-4}\).
\begin{center}
    \includegraphics[scale=0.67]{a2c_out}
\end{center}
The graph shows that the moving average reward converges within only 50 
iterations, that is, 250 episode, 75000 parameter updates for each network, 
and stays quite stable afterwards.

\section{Actor-Critic, Reinforce, and PPO}
\subsection{Algorithms}
To further show how Actor-Critic algorithm performs, we compare it to another
two popular algorithms: Reinforce and PPO. Below are pseudocodes for the two
algorithms:
\begin{algorithm}[H]
\caption{Reinforce for our financial problem}
\begin{algorithmic}[1]
\State Initialize ANN policy parameterization 
\(\pi\left(c\middle|s, \bm{\theta}\right)\), with any \(\bm{\theta}\)
\State Step sizes \(\alpha_{\bm{\theta}}>0\)
\Loop \hspace{0.3mm} (for each episode)
    \State Reset the environment
    \State Generate an episode \(S_0, C_0, R_1, ..., S_{T-1}, C_{T-1}, R_T\),
    following \(\pi\left(\cdot\middle|\cdot, \bm{\theta}\right)\)
    \For{\(t=0, 1, ..., T\)}
        \State \(G \gets \sum_{k=t+1}^{T} \gamma^{k-t-1}R_k\)
        \State \(\bm{\theta} \gets \bm{\theta} + \alpha_{\bm{\theta}}\gamma^t
        G\nabla{\ln{\pi\left(C_t\middle|S_t, \bm{\theta}\right)}}\)
    \EndFor
\EndLoop
\end{algorithmic}
\end{algorithm}

\begin{algorithm}[H]
\caption{PPO for our financial problem}
\begin{algorithmic}[1]
\State Initialize ANN policy parameterization 
\(\pi\left(c\middle|s, \bm{\theta}\right)\), with any \(\bm{\theta}\)
\State Initialize ANN state-value function parameterization
\(\hat{v}(s,\bm{w})\) with any \(\bm{w}\).
\State Step sizes \(\alpha_{\bm{\theta}}>0\), \(\alpha_{\bm{w}}>0\)
\For{iteration \(k=0, 1, 2, ...\)}
    \State Collect set of trajectories \(\mathcal{D}_k=\{\tau_i\}\), 
    following \(\pi\left(\cdot\middle|\cdot, \bm{\theta}_k\right)\), where 
    \(\tau = S_0, C_0, R_1, ..., S_{T-1}, C_{T-1}, R_T\)
    \For{\(t=0, 1, ..., T\) in each trajectory \(\tau\) in \(\mathcal{D}_k\)}
        \State \(G_t \gets \sum_{k=t+1}^{T} \gamma^{k-t-1}R_k\)
        \State \(A_t \gets G_t - \hat{v}(S_t)\)
    \EndFor
    \Loop \hspace{0.3mm} (update multiple times)
    \State \(\bm{\theta} \gets \underset{\bm{\theta}}{\text{arg\:max}}
    \frac{1}{|\mathcal{D}_k|T}\sum_{\tau\in\mathcal{D}_k}^{} \sum_{t=0}^{T} 
    \min{\left(
            \frac{\pi\left(c_t\middle|S_t, \bm{\theta}\right)}
            {\pi\left(c_t\middle|S_t, \bm{\theta}_k\right)}A_t,
            \text{clip}\left(
                \frac{\pi\left(c_t\middle|S_t, \bm{\theta}\right)}
                {\pi\left(c_t\middle|S_t, \bm{\theta}_k\right)},
                1-\epsilon, 1+\epsilon
            \right)A_t
    \right)}\)
    \State \(\bm{w} \gets \underset{\bm{w}}{\text{arg\:min}}
    \frac{1}{|\mathcal{D}_k|T}\sum_{\tau\in\mathcal{D}_k}^{} \sum_{t=0}^{T} 
    \Bigl(\hat{v}(S_t, \bm{w})-G_t\Bigr)^2\)
    \EndLoop
\EndFor
\end{algorithmic}
\end{algorithm}

\subsection{Training result}
To make better comparison, the policy function of the two algorithms
is implemented the same way as we did in Actor-Critic algorithm, and the same 
is true for the value function used in PPO method. 
\vspace{3mm}\newline The following graph shows the training result of these 
three algorithms (300 moving average reward vs iteration), with 600 iterations
(5 episodes per iteration), \(T=300\), \(\gamma=0.99\), 
\(\alpha_\theta=\num{3e-4}\), \(\alpha_{\bm{w}}=\num{3e-10}\) (if applicable),
\(\epsilon=0.5\) (if applicable).
\begin{center}
    \includegraphics[scale=0.67]{out}
\end{center}
As is shown in the graph, the Actor-Critic algorithm converges the fastest,
within 50 iterations, followed by Reinforce algorithm, which converges within
about 250 iterations, despite a small drop of moving average around iteration 
300, which is quickly corrected by the algorithm within 50 iterations. PPO 
algorithm converges the slowest, with about 300 iterations. Despite of the 
speed of convergence, the moving average rewards of convergence for these 
three algorithms are nearly the same (around 400), this is as expected.

\end{document}